Efficient Calculation of Polynomial and Interaction Features on Sparse Matrices

### 2015-09-21 paper/paper2.tex
need john's proof

which results in $\binom{D+1}{D-1} = \frac{D^2+D}{2}$ additional features, and so the generation of 
these features for a matrix of size $N \times D$ has a time complexity of $O(ND^2)$. In the general case of degree $k$, there 
are $\binom{D+k-1}{D-1}$ polynomial features and the time complexity is $O(ND^k)$.

@@

The first of these necessary components is gotten simply by using the appropriate data structure
to store the sparse matrix (e.g. storing it in compressed sparse row format). The second component 
is not only less readily available, but its necessity is less obvious. To make this need more clear, consider the
algorithm for the brute force calculation of polynomial features:

@@

This work also gives a general pattern for calculating polynomial features of arbitrary degrees:
instead of finding a mapping from the 2-tuples of indices onto the space $\{0, 1, ..., \frac{D^2+D}{2}-1\}$,
construct a mapping from k-tuples of indices onto the space $\{0, 1, ..., \binom{D+k-1}{D-1}-1\}$, where $k$ is the degree of the polynomial.
In terms of the upper triangular matrix analogy, this corresponds with mapping the upper k-simplex of a $(D, D, ..., D)$-tensor onto
the space $\{0, 1, ..., \binom{D+k-1}{D-1}-1\}$. For the average case of $dD$ nonzero elements per row, 
there will be $\binom{dD+k-1}{dD-1} = $ polynomial features, so the average time complexity for an $N \times D$ matrix
is $\Theta(Nd^kD^k)$, which means that its time complexity polynomially decreases with respect to the density
compared to the dense algorithm (which has $O(ND^k)$ and $\Theta(ND^3)$).

>>>>

which results in $\binom{D+1}{D-1} = \frac{D^2+D}{2}$ additional features, so the generation of 
these features for a matrix of size $N \times D$ has a time complexity of $O(ND^2)$. In the general case of degree $k$, there 
are $\binom{D+k-1}{D-1}$ polynomial features and the time complexity is $O(ND^k)$.

@@

The first of these necessary components is obtained simply by using the appropriate data structure
to store the sparse matrix (e.g. storing it in compressed sparse row form). The second component 
is not only less readily available, but its necessity is less obvious. To make this need more clear, consider the
algorithm for the brute force calculation of polynomial features:

@@

This work also gives a general pattern for calculating polynomial features of arbitrary degrees:
instead of finding a mapping from the 2-tuples of indices onto the space $\{0, 1, ..., \frac{D^2+D}{2}-1\}$,
construct a mapping from k-tuples of indices onto the space $\{0, 1, ..., \binom{D+k-1}{D-1}-1\}$, where $k$ is the degree of the polynomial.
In terms of the upper triangular matrix analogy, this corresponds with mapping the upper k-simplex of a $(D, D, ..., D)$-tensor onto
the space $\{0, 1, ..., \binom{D+k-1}{D-1}-1\}$. For the average case of $dD$ nonzero elements per row, 
there will be $\binom{dD+k-1}{dD-1} = $ polynomial features, so the average time complexity for an $N \times D$ matrix
is $\Theta(Nd^kD^k)$, which means that its time complexity polynomially decreases with respect to the density
compared to the dense algorithm (which has $O(ND^k)$ and $\Theta(ND^k)$).
